### 2021-01-10 paper/main.tex
Initial commit of in-progress NSR20 paper files

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% Additional Packages:
\usepackage{graphicx}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage{textcomp} % For \textdegree (also try \usepackage{gensymb} which gives \degree)
\usepackage{soul} % For highlighting text

% Macro Definitions:
\def\s{\sigma}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\title{Automated data-driven symbolic modeling of the lift forces acting on flapping wings}

\author[1,*]{Charles Richter}
\author[2,3]{Hod Lipson}
% \author[1,2,+]{Christine Author}
% \author[2,+]{Derek Author}
\affil[1]{Cornell University, Sibley School of Mechanical and Aerospace
Engineering, Ithaca, 14850, United States}
\affil[2]{Columbia University, Department of Mechanical Engineering, New York,
10027, United States}
\affil[3]{Columbia University, Data Science Institute, New York, 10027, United
States}

\affil[*]{car45@cornell.edu}

% \affil[+]{these authors contributed equally to this work}

\keywords{flapping flight, unsteady aerodynamics, data mining, symbolic
regression, evolutionary computation}

\begin{abstract}
We explore a data-driven approach to symbolically modeling the lift produced by
the flapping wing of a micro air vehicle.  The aerodynamics of flapping flight
pose a modeling challenge due to complex effects related to vortex dynamics,
wake interaction, aeroelasticity and three-dimensional flow, yet simple and
accurate equations are useful for engineering purposes and optimization of wing
design.  Experiments were conducted to measure the flapping kinematics and
instantaneous lift forces produced by eleven different 3D printed wing
geometries at three different flapping amplitudes and multiple speeds, totaling
68 distinct sets of data.  58 of these data sets were used to train data-driven
models using a symbolic regression algorithm, while the remaining sets were
reserved for model testing.  Data-driven models were found to be more accurate
and much simpler than existing quasi-steady analytical models consistent with
aerodynamic theory.  Furthermore, data-driven models were found to separate
distinct physical effects into inertial and aerodynamic components bearing
remarkable resemblance to their analytical counterparts. This result suggests
that a data-driven approach could be a powerful tool in separating and
identifying the structural building blocks of a wide variety of unknown
dynamical systems.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

% \noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.

\section*{Introduction}

% The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

Forces produced by the flapping wings of insects, birds and micro air vehicles
are of crucial importance for flight, yet they are subtle and difficult to
model.  Many studies have developed equations and solution techniques to
understand the fluid flow around these wings, to compute control actions
required for stable flight, or to optimize wing design. Aerodynamic studies
related to insect flight and intermediate Reynolds numbers (Re 100-10,000) have
generally focused on either numerical solutions of Navier-Stokes equations in
two or three dimensions \cite{wu2004unsteady,wang2000two}, quasi-steady
approximations of the observed behavior
\cite{sane2002aerodynamic,ellington1984aerodynamics}, or a comparison of both
methods \cite{pesavento2004falling}.  The Navier-Stokes equations offer
important insights into the full character of the fluid around the wing
including vorticity and fluid-structure interaction, but are complex and
computationally expensive to calculate.  Quasi-steady approximations, in
contrast, capture most of the character of the aerodynamic forces without any
detailed modeling of the fluid flow.  These equations are phenomenological in
nature and constitute an assembly of terms that are consistent with aerodynamic
theory that can be used to represent lift, drag and other forces, as functions
of a wing's kinematic variables such as velocity and angle of attack.  Both
methods have shown qualitative and quantitative agreement with experimental
measurements \cite{wang2004unsteady}.

% \subsection{Automated modeling}

Quasi-steady representations of lift are particularly well suited to the process of symbolic
regression, which harnesses computational power to conduct an automated search over the space of all
kinematic variables and geometric parameters to find equations that most accurately predict the
forces on a wing.  Symbolic regression algorithms are based on evolutionary computation
\cite{koza1992genetic,forrest1993genetic} and have been used to model complex systems defined by explicit
\cite{duffy2002using,fillon2007symbolic,bautu2005symbolic}, implicit \cite{schmidt2010symbolic}, iterated
\cite{schmidt2009solving} and differential equations \cite{bongard2007automated}.  The
process underlying this evolutionary computation approach is to generate an initially random
population of equations combining time-series data and constant variables with algebraic operators,
analytic functions and constants.  These candidate equations are then evaluated to measure their
accuracy in predicting the measured output of the system or partial derivatives within the data.  In
each iteration of the algorithm, the best candidates are retained and varied by recombining and
varying their subexpressions. The algorithm is terminated when a desired level of accuracy is
reached.  In this case, the time-series data includes the orientations, velocities and accelerations
of wings, the constant variables include chord lengths, wingspans, centroids of wing area and other
physical constants, and the known output of the system is the measured vertical force produced
during flapping.  The resulting equations returned by the algorithm are quasi-steady models that can
be compared to existing equations from the literature.

% \subsection{Measurement of lift forces}

Numerous studies have measured the forces produced by a flapping wing in
various experimental setups including flapping in a liquid bath such as oil
\cite{walker2009photogrammetric}, by tracking the kinematics of insects in
flight \cite{fry2003aerodynamics,ristroph2009automated} and by tracking the
detailed wing motions and visualizing the airflow around the wings of insects
in a wind tunnel \cite{young2009details}.  Only recently have studies directly
measured the forces involved in flapping with at-scale robotic wings or actual
insects in air \cite{whitney2010aeromechanics,graetzel2008real}.  In general,
measuring the forces produced by flapping wings in air is difficult because the
magnitude of the forces is small, requiring custom-built capacitive or MEMS
force sensors, and because the speed of flapping and prominence of inertial
effects cause significant vibrations that obscure the force measurement.  The
flapping wings examined in this work represent a size range appropriate for
micro air vehicles, and produced forces large enough to be measured by a
commercially available load cell.

Here we present a series of experiments measuring the kinematics and lift forces produced by eleven
differently sized wings flapping at multiple speeds and amplitudes with Reynolds numbers between
2500 and 5000. Several quasi-steady models from the literature are shown to exhibit strong agreement
with the experimental data.  Then a set of data-driven equations were automatically generated from
the data using the symbolic regression algorithm developed by Schmidt and Lipson \cite{schmidt2009distilling}.
These data-driven models exhibit several remarkable properties.  First, they are at least as
accurate as quasi-steady equations from the literature yet they are much simpler.  Second, they
naturally separate distinct, understandable physical effects including terms corresponding to
inertial forces and aerodynamic lift.  Finally, it is shown that models generated from data of a
single experiment do not have the property of separating distinct physical effects, and they fail to
accurately predict forces produced by a wing of a different geometry or flapping pattern.  It is
only when models are trained on a large set of experiments including many different wings and
flapping patterns that over-fit candidate equations are discarded and those that reflect the true
underlying physics are maintained.  The models trained on a diverse set of data are shown to
accurately predict the forces produced by wings of similar size and flapping kinematics that were
not included in the training set. The ability of the symbolic regression process to infer distinct
physical contributions to to the total lift force based on data alone suggests it may be a powerful
tool in identifying the underlying structure of a wide variety of complex systems.

\section*{Results}

% Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
%
% \subsection*{Subsection}
%
% Example text under a subsection. Bulleted lists may be used where appropriate, e.g.
%
% \begin{itemize}
% \item First item
% \item Second item
% \end{itemize}
%
% \subsubsection*{Third-level section}
%  
% Topical subheadings are allowed.

\subsection{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate equations on 58 of the 68 total experiments simultaneously, with the remaining ten reserved for testing.  These 58 training sets were intended to represent a diverse sampling of wing geometry, flapping speed and amplitude combinations.  The results from these 58 experiments were compiled into a single dataset listing roughly two periods of smoothed, registered flapping data per experiment.  The data consisted of instantaneous measured vertical force along with the measured value of each kinematic variable, the first and second derivatives of those variables, and the geometric parameters necessary to describe the wing's shape and distribution of area, totaling 31 separate variables.

By training on 58 experiments simultaneously, the symbolic regression algorithm generated equations that fit all of the data, using the some, but not all, of the provided variables to account for variations in kinematics and wing geometry between experiments.   Table~\ref{table:eqmodels} lists five of the leading candidate models, which were trained using mean absolute error (MAE) as the fitness metric during the search. The search was conducted for approximately 24 hours, using 60 cores on a computing cluster.

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & {\bf MAE (N)}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.46\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.28\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.28\e{-3}\\
Whitney-Wood & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.21\e{-3}\\
Pesavento-Wang & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical equations of lift.
The models are sorted by equation size, which is calculated as the sum of arithmetic and
trigonometric operators required to compute a force prediction from kinematic and geometric data.
MAE reports average mean absolute error (in Newtons of vertical force) for the ten test
experiments.}
\end{table}

\subsection{Analytical models for comparison}
The analytical models used for comparison are also listed in Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces added to a collection of aerodynamic terms.  The primary aerodynamic contribution in all quasi-steady models is a translational lift term proportional to the squared velocity of each span-element of the wing that is related to the wing's angle of attack and empirically determined coefficient of lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when out-of-plane motions are included. This translational lift term is given by:
\begin{equation}
L_{trans}=(\nicefrac{1}{2})\rho v^2AC_L,
\end{equation}
where \(v\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.  The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
\begin{equation}
C_L = C_{Lmax}\sin(2\alpha).
\end{equation}

Other aerodynamic contributors include rotational lift proportional to the product of a wing's rotational and translational velocities, labeled $L_{rot}$, as well as "added-mass" effects related to the accelerations of fluid surrounding the wing \cite{sedov1965two}.  These added mass terms are proportional to the wing's linear acceleration ($AM_{trans}$), rotational acceleration ($AM_{rot}$), or a product of the two ($AM_{trans-rot}$).  Two sets of quasi-steady equations from the literature include various combinations of such terms \cite{whitney2010aeromechanics,pesavento2004falling}, which allow more flexible model fitting through additional coefficients.

Inertial forces acting on the wings are modeled as a sum of two contributions.  The primary contribution results from the pendulum-like swinging motion that occurs when the deflection angle ($\psi$) flips at stroke reversal.  The secondary contribution results from small angle deviations from the horizontal flapping plane ($\theta$).  These angles are depicted in Figure~\ref{fig:apparatus}.  These inertial forces are given by:
\begin{equation}\label{sdbla}
F_{pendulum} = Y_{CM}M(\ddot{\psi}\sin\psi + \dot{\psi}^2\cos\psi)
\end{equation}
and
\begin{equation}\label{sdbla2}
F_{deviation} = -X_{CM}M(\ddot{\theta}\cos\theta - \dot{\theta}^2\sin\theta),
\end{equation}
\noindent where \(X_{CM}\) and \(Y_{CM}\) are the coordinates of the center of wing area and \(M\) is the mass of the wing surface.

\begin{figure}%[!htb]
\centering
\begin{subfigure}{0.3\textwidth}
\includegraphics[trim=0 15 0 5, clip, width=\textwidth]{figures/apparatus}
\caption{\label{fig:apparatus}}
\end{subfigure}
\qquad
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/pareto_mae_nofit_PNAS}
\caption{\label{fig:pareto_mae}}
\end{subfigure}
\caption{\subref{fig:apparatus} Schematic diagram of the experimental apparatus showing three
  kinematic variables of the wing motion: stroke ($\phi$), deflection ($\psi$) and deviation
  ($\theta$). \subref{fig:pareto_mae} Pareto plot of all data-driven (``EQ Models'') and analytical
  models using mean values for model coefficients. All data-driven models were smaller and more
  accurate than analytical models.
}
\end{figure}

\subsection{Evaluation of Model Performance}
An evaluation of model performance allows comparison between analytical models and the equations developed through the data mining process.  Equations are judged by their accuracy and complexity.  In this case, the complexity or size of an equation was measured by simply counting the number of arithmetic and trigonometric operators used to combine the measured kinematic variables and geometric properties into an equation for instantaneous vertical force.  Both our data-driven models and the analytical equations are measured in this way.  For fairness, constants in the analytical models that would have been lumped together into larger coefficients by the symbolic regression algorithm (such as material density, air density, \(\pi\) or factors of integration) are also lumped together in the calculations of size for the analytical equations.  The size of the inertial calculations was added to the sizes of analytical lift calculations.  The 
pendulum motion has an equation size of 13 and the total inertial calculation including out-of-plane deviation has a size of 27.

Table~\ref{table:eqmodels} shows the size of all equations considered in this study, and indicates that the analytical models are all significantly more complex than our data-driven models.  One of the reasons for high complexity in the analytical models is the dependence on wing geometry and integration of forces over the span and chord of the wing.  A second reason for the complexity in analytical models is the number of geometric and coordinate transformations required to compute the necessary linear and rotational velocities.  In this sense, several terms generated during symbolic regression can be considered to be approximations of much more subtle calculations of kinematic quantities.

The accuracy of each equation was measured as the mean absolute error (MAE) between the predicted and actual vertical force (in Newtons) produced by the flapping wing.  Figure~\ref{fig:pareto_mae} shows the Pareto plot of all models considered, depicting accuracy as a function of model complexity. It is important to note that all data-driven models have between three and five coefficients, so tuning these coefficients allows further minimization of error compared to the analytical models, which have between one and three.  Nevertheless, discovery of models with smaller error than theoretical predictions that are simultaneously much simpler is a significant feat for the data mining process and symbolic regression.

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_PNAS}
\caption{Comparison of force predictions between $EQ_4$ (blue) and Equation~\eqref{eq:badeq} (red).
  $EQ_4$ was trained on 58 distinct experiments simultaneously, including the data shown in (A),
  whereas Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown in (A).  Panel (A)
  shows that both models perform well on this data within their training sets.  However, when
  presented with data outside their training set (B,C,D), only $EQ_4$ is able to generalize and
  provide accurate predictions, while Equation~\eqref{eq:badeq} fails to do so.}
\label{fig:eq_goodbad}
\end{figure}

\subsection{Performance of models trained on one experiment versus many experiments}
Figure~\ref{fig:eq_goodbad}(A) illustrates the predictive performance of two models on data within their training sets, while Figure~\ref{fig:eq_goodbad}(B,C,D) illustrate the performance of those same two models on data outside their training sets. The differences between these two models result from the fact that one of them, $EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained only on the single experiment shown in Figure~\ref{fig:eq_goodbad}(A):
\begin{align}
 Force = \frac{C_1 + C_2\psi^2 - C_3\psi}{C_4 + \ddot{\theta} + C_5\dot{\theta} + \cos(C_6 -
  \frac{C_7}{\psi^2} - C_8\psi)} - C_9.
\label{eq:badeq}
\end{align}
Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not over-fit but in fact generalizes well to new data. Equation~\eqref{eq:badeq}, on the other hand, fails to accurately predict the recorded data outside its training set.  Insufficient data and over-fitting are potential pitfalls of data-driven modeling, but are overcome in this case with a diverse sampling of training data.

Since Equation~\eqref{eq:badeq} is unable to accurately predict forces outside its training set, it is unlikely to offer any insight into the underlying physical processes.  $EQ_4$, however, must capture some of the underlying structure of the physics in order to generalize to data upon which it was not trained.  In data-driven modeling, training on many experiments simultaneously helps to reject candidate equations that may fit some, but not all, of the data, and helps to retain those models that reflect the underlying physics and offer insight into the structure of the additive components that compose the observed behavior of a complex dynamical system.

\subsection{Correspondence of physical terms}
One powerful result of the data-driven modeling approach is that in many cases, certain terms from the data-driven models closely match physically understandable terms from the analytical models.  The correspondence between these terms is listed in Table~\ref{table:eqcorrespondence}. Taking model $EQ_4$ as an example, all three terms show strong resemblance to inertial and aerodynamic components of the analytical models.  Figure~\ref{fig:eqcomparison} shows a comparison between the simple velocity-squared aerodynamic lift prediction and the second term of $EQ_4$: \(C_2RX_{CM}\psi\dot{\phi}\).  These two calculations are numerically very similar, yet the third term of $EQ_4$ is much simpler than $L_{trans}$.

Similarly, the sum of terms 1 and 3, \(C_1RY_{CM}\dot{\psi}^2\cos(\psi) - C_3R\ddot{\theta}\), closely approximates the calculation of inertial forces including the pendulum motion of the wing as well as the out-of-plane deviation.  Figure~\ref{fig:eqcomparison} also shows these two calculations, indicating very similar structure.  While $EQ_4$'s approximation of these inertial forces does not precisely match the analytical calculation, it follows a very similar pattern and is much simpler.  Remarkably, these inertial terms utilize some of the same variable combinations as the analytical calculations to express the inertial force, namely \(Y_{CM}\dot{\psi}^2\cos(\psi)\) and \(\ddot{\theta}\), with different coefficients to express the geometric variation between wings and small angle approximations \(\cos\theta\approx1\), \(\sin\theta\approx0\) for small out-of-plane deviation angles.  Since \(|\theta|<2^{\circ}\) in general, the small angle approximation is valid.

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|l|}
\hline
{\bf Contribution} & {\bf Analytical} & {\bf Data-Driven ($EQ_4$)}\\
\hline
Aerodynamic Lift & \(\nicefrac{1}{2}\rho v^2AC_{Lmax}\sin(2\alpha)\) & \(C_2RX_{CM}\psi\dot{\phi}\)\\
Pendulum 1 & \(M\cdot Y_{CM}\ddot{\psi}\sin{\psi}\) & \it{None} \\
Pendulum 2 & \(M\cdot Y_{CM}\dot{\psi}^2\cos{\psi}\) & \(C_1R\cdot Y_{CM}\dot{\psi}^2\cos{\psi}\) \\
Deviation 1 & \(\approx-MX_{CM}\ddot{\theta}\) & \(- C_3R\ddot{\theta}\)\\
Deviation 2 & \(\approx0\) & \it{None} \\
%Deviation 2 & \(X_{CM}M\dot{\theta}^2\sin\theta\) & \(<none>\)\\
\hline
\end{tabular}
\caption{\label{table:eqcorrespondence}Correspondence between analytical and data-driven terms.}
\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=.48\textwidth]{figures/eq_comparison_PNAS}
\caption{Comparison between terms of the analytical expressions (red) and data-driven models (blue).
  (A) Primary translational aerodynamic lift proportional to the square of the wing velocity. (B)
  Inertial forces related to the pendulum motion of the wing and deviation out of the stroke plane.
  The modeling process automatically decoupled aerodynamic and inertial effects with no prior
  knowledge of the system.  The total measured force is shown for reference (black).}
\label{fig:eqcomparison}
\end{figure}

\section*{Discussion}

% The Discussion should be succinct and must not contain subheadings.

Table~\ref{table:eqcorrespondence} lists the corresponding terms between the
basic analytical expressions and the data-driven model $EQ_4$.  While two of
the inertial terms share the same combinations of variables, the data-driven
expression for aerodynamic lift is very different from its analytical
counterpart, and one portion of the pendulum motion is omitted altogether from
the data-driven models.  The reasons for these differences lie in the nature of
the symbolic regression algorithm, which is designed to return only the most
parsimonious expressions capable of predicting the experimental data at a given
level of accuracy.  Therefore, key physical effects may be represented
numerically by terms that are symbolically different and often much simpler
than their analytical counterparts.  This behavior of automatically separating
distinct dynamical building blocks is one of the most powerful aspects of
symbolic regression.  By sweeping the range of several kinematic and geometric
parameters in the 58 sets of training data, any candidate models that
incorrectly mix or couple two separate physical effects are discarded upon
testing on a data set with different parameters.

Nevertheless, in some domains such as flapping flight, there exist high degrees
of correlation between certain variables.  For example, the center of area
location in a rectangular wing is closely related to its span and chord length.
Similarly, kinematic variables such as stroke and deflection angles are highly
correlated with each other and their derivatives across all 58 experiments,
making these variables approximately interchangeable with each other when
coefficients are adjusted accordingly.  This behavior provides an explanation
for why \(\psi\dot{\phi}\) is used as an approximation for an aerodynamic lift
term proportional to \(v^2\).  These correlations between variables are
inherent in highly symmetric, periodic phenomena such as flapping flight.  In
principle, additional experiments could be performed on flapping surfaces that
are not bound to the natural kinematics of flapping wings to explore the full
parameter space and disambiguate non-physical terms in the candidate models.

% \section{Conclusions} 
The major thrust of this work was to apply a data-driven approach to modeling
the forces on a flapping wing.  The models produced by our symbolic regression
algorithm were shown to be more accurate than any of the analytical models, and
were much simpler, in some cases by an order of magnitude in equation size.
Furthermore, by dissecting these equations and examining individual terms, it
was observed that the data-driven models separated the measured forces into
understandable physical effects.  A lift term was found that was numerically
similar to the standard translational lift, and an inertial term was found that
was numerically similar to the complete calculation of the pendulum and
out-of-plane motions with some analytical terms in common.  Finally, by
training data-driven models on experimental data from 58 individual
wing-amplitude-speed combinations simultaneously, the symbolic-regression
algorithm was able to discard models that did not generalize, while retaining
those whose terms captured the underlying physical processes.

This application of a data-driven approach suggests that symbolic regression
could be useful in phenomenological modeling and guiding the exploration of a
wide variety of complex systems.  Its ability to automatically identify
distinct physical effects that are consistent across all datasets makes it a
powerful tool for identifying the structure of contributions from building
blocks that make up the complete observed behavior of a dynamical system.

\section*{Methods}

% Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.

\subsection{3D Printed wings and actuation mechanism}
The experimental wings were fabricated using an Objet Connex500 3D printer with
the rigid Objet FullCure720 material.
Flapping flight with 3D printed wings was reported in \cite{richter2011untethered}. The
wing design consisted of an interlocking root and tip, allowing the wing
surface to pivot naturally during flapping from its neutral vertical position
to an angle of attack of approximately 45\textdegree. The wing design is illustrated in
Figure~\ref{fig:wing1} and the hinge design is
illustrated in Figure~\ref{fig:root2}.  In total, eleven wings were used,
ranging in wingspan from 40 mm to 120 mm and ranging in chord length from 40 mm
to 80 mm. Flapping amplitudes were constrained by the geometry of the wing root
design to be set at 72\textdegree, 81\textdegree, and 96\textdegree, and each
wing was flapped at speeds ranging between 1 Hz and 3.5 Hz for each of these
amplitude settings.  68 sets of data were recorded, representing Reynolds
numbers between 2500 and 5000.

% See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
\begin{figure}[h]%[!htb]
\centering
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/wing1}
    \caption{\label{fig:wing1}}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[trim=0 105 0 0, clip, width=0.95\textwidth]{figures/root2}
    \caption{\label{fig:root2}}
\end{subfigure}
\caption{Detail of 3D printed wing \subref{fig:wing1} and wing root \subref{fig:root2} design
  showing concentric-cylindrical design to allow passive pivoting of the wing surface.}
\end{figure}

\subsection{Measurement apparatus}
The apparatus used to measure the forces and kinematics of flapping wings was
based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
logging.  This force measurement system was capable of a sampling rate of 1 kHz
and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
a crank and connecting rod used to drive the experimental wings in a roughly
sinusoidal flapping motion. This system is pictured in
Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency
of the load cell and wing driving mechanism was roughly 70 Hz, the natural
frequency for the first bending mode of the experimental wing was approximately
26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
All experimental force measurements were filtered using a high-order zero-phase
digital low-pass filter with a cutoff frequency of 7.5 times the flapping
frequency, which preserved the primary inertial and aerodynamic forces acting
on the wings while eliminating the faster vibrations of the wing structure and
the load cell apparatus.

% % See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
\begin{figure}%[!htb]
\centering
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[height=5cm]{figures/apparatus2}
    \caption{\label{fig:apparatus2}}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[height=5cm]{figures/apparatus3}
    \caption{\label{fig:apparatus3}}
\end{subfigure}
\caption{ \subref{fig:apparatus2} Front view of the flapping wing apparatus showing the wing mounted
  on its driving mechanism attached to the force sensor. \subref{fig:apparatus3} Close-up view of
  the 3D printed frame and crank mechanism used to drive the flapping motion of the wing.}
\end{figure}

\subsection{High-speed measurement of flapping kinematics}
A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
210 frames per second from a top-down perspective as well as a side view
provided by a mirror mounted in the field of view.  The main beam of each wing
was colored red to enable precise tracking of the stroke angle.  The top-down
projected area of the wing was used to extract deflection angle, and tracking
of the wing tip in the side-view was used to extract the angle of deviation out
of the wing's stroke plane.  These three angles precisely define the
configuration of the wing.  A schematic diagram of this apparatus is shown in
Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
to extract the kinematics of the wing motion from high-speed video.  The
red-colored main beam was identified with a color filter and a line was fitted
using a Hough Transform to identify the stroke angle.  The wing deflection
angle was computed by measuring the projected area of the wing in the image and
compared with the known dimensions of the rectangular wing shape. The
out-of-plane deviation angle of the wing was measured by tracking the leading
and trailing edge corners of the wingtip in the side-view of the mirror. We refer the reader to our
previous work for a more detailed description of our methods of extracting wing kinematics from
high speed video and processing the measured data\cite{richter2012thesis}.
Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
experiment.

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{figures/stream}
% \caption{Legend (350 words max). Example legend text.}
% \label{fig:stream}
% \end{figure}

\begin{figure}
\centering
\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regforce1}
\caption{Measured vertical force.}
\label{fig:regforce}
\end{subfigure}
\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regstroke1}
\caption{Wing stroke angle.}
\label{fig:regstroke}
\end{subfigure}

\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regdef1}
\caption{Wing deflection angle.}
\label{fig:regdef}
\end{subfigure}
\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regdev1}
\caption{Wing deviation angle.}
\label{fig:regdev}
\end{subfigure}
\caption{Measured \subref{fig:regforce} vertical force, \subref{fig:regstroke} wing stroke angle,
  \subref{fig:regdef} deflection angle, \subref{fig:regdev} and deviation angle from one experiment,
  registered and averaged over 18 flapping periods.}
\label{fig:kinematics}
\end{figure}

\bibliography{main}

% \noindent LaTeX formats citations and references automatically using the bibliography records in your .bib file, which you can edit via the project menu. Use the cite command for an inline citation, e.g.  \cite{Hao:gidmaps:2014}.
%
% For data citations of datasets uploaded to e.g. \emph{figshare}, please use the \verb|howpublished| option in the bib entry to specify the platform and the link, as in the \verb|Hao:gidmaps:2014| example in the sample bibliography file.

\section*{Acknowledgements}

% Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

This research was funded by U.S. National Science Foundation (NSF) Grant ECCS 0941561 on
Cyberenabled Discovery and Innovation (CDI).

\section*{Author contributions statement}

% Must include all authors, identified by initials, for example:
% A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

C.R. and H.L. conceived the experiments, C.R. conducted the experiments, and C.R. and H.L. analyzed
the results. Both authors reviewed the manuscript.

% \section*{Additional information}
%
% To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement). 
%
% The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on behalf of all authors of the paper. This statement must be included in the submitted article file.

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{figures/stream}
% \caption{Legend (350 words max). Example legend text.}
% \label{fig:stream}
% \end{figure}

% \begin{table}[ht]
% \centering
% \begin{tabular}{|l|l|l|}
% \hline
% Condition & n & p \\
% \hline
% A & 5 & 0.1 \\
% \hline
% B & 10 & 0.01 \\
% \hline
% \end{tabular}
% \caption{\label{tab:example}Legend (350 words max). Example legend text.}
% \end{table}

% Figures and tables can be referenced in LaTeX using the ref command, e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}